---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)      # data wrangling
library(naniar)         # missing values visualization
library(visdat)         # missing values visualization
library(ComplexHeatmap) # missing values visualization


training <- read_csv("train.csv")
testing <- read_csv("test.csv")
```



## Missing data
I start by fixing the obvious missing values. The only problematic variable seems to be year built for garage where the garace is missing. One way would be to replace NA with year built for the house itselfe.   
```{r}
# Missing data according to data_description.txt file

training$PoolQC[is.na(training$PoolQC)] <- "no_pool"
training$MiscFeature[is.na(training$MiscFeature)] <- "none"
training$Alley[is.na(training$Alley)] <- "no_alley"
training$Fence[is.na(training$Fence)] <- "no_fence"
training$FireplaceQu[is.na(training$FireplaceQu)] <- "no_fireplace"

#  Basement
training$BsmtFinType2[is.na(training$BsmtFinType2)] <- "no_basement"
training$BsmtFinType1[is.na(training$BsmtFinType1)] <- "no_basement"
training$BsmtExposure[is.na(training$BsmtExposure)] <- "no_basement"
training$BsmtQual[is.na(training$BsmtQual)] <- "no_basement"
training$BsmtCond[is.na(training$BsmtCond)] <- "no_basement"

# Garage
training$GarageType[is.na(training$GarageType)] <- "no_garage"
training$GarageFinish[is.na(training$GarageFinish)] <- "no_garage"
training$GarageQual[is.na(training$GarageQual)] <- "no_garage"
training$GarageCond[is.na(training$GarageCond)] <- "no_garage"


df_missing <- map_df(training, function(x) mean(is.na(x))) %>%
        gather()

ggplot(filter(df_missing, value > 0),
       aes(x = fct_reorder(key, value),
           y = value)) +
        geom_bar(stat = "identity") +
        coord_flip()

```


### Further investigation into missing values
```{r}
convert_missing <- function(x) ifelse(is.na(x), 0, 1)

house_missing <- apply(training, 2, convert_missing)

Heatmap(house_missing,
        name = "Missing",
        column_title = "Predictors",
        row_title = "Samles",
        col = c("black", "lightgrey"),
        show_heatmap_legend = FALSE,
        row_names_gp = gpar(fontsize = 0)) # text size for row names


gg_miss_upset(training)


```


### LotFrontage
LotFrontage referce to *Linear feet of street connected to property*. When compared to SalePrice there doesn't seems to be that much of a difference in sale price whether or not the LotFrontage is missing. It could be that the actual value is 0 but we dont' know for sure. So we can impute the missing values using KNN or Bagging and compare it to imputing the missing values with zero.  

#### Numeric variables and missingness of LotFrontage
No clear pattern.  
```{r}

df_numeric <- training %>% 
        select_if(is.numeric) %>% 
        mutate(LotFrontage_cat = ifelse(is.na(LotFrontage), "missing", "no_missing")) %>% 
        select(-Id,- LotFrontage) %>% 
        select(LotFrontage_cat, everything()) %>% 
        gather("key", "value", 2:37)

ggplot(df_numeric,
       aes(x = LotFrontage_cat,
           y = value,
           col = LotFrontage_cat)) + 
        geom_violin() +
        facet_wrap(~key, ncol = 6, scales = "free_y")

```

#### Categorical variables and connection to LotFrontage

```{r}
df_categorical <- training %>% 
        select_if(is.character)

df_lot <- training %>% 
        select(Id, LotFrontage)

df_categorical <- bind_cols(df_lot, df_categorical) %>% select(-Id)

df_categorical <- df_categorical %>% 
        gather("key", "value", 2:44) %>% 
        group_by(key, value) %>% 
        summarise(missing = mean(is.na(LotFrontage)))


ggplot(df_categorical,
       aes(x = value,
           y = missing,
           fill = value)) +
        geom_bar(stat = "identity") +
        facet_wrap(~key)

```


```{r}

lot_miss <- training %>% 
        mutate(LotFrontage_miss = ifelse(is.na(LotFrontage), "missing", "not_missing"))


ggplot(lot_miss,
       aes(x = LotFrontage_miss,
           y = SalePrice)) + 
        geom_violin()

ggplot(lot_miss,
       aes(x = LotFrontage_miss,
           y = SalePrice)) + 
        geom_boxplot()



```

