---
title: "House Prices: Advanced Regression Techniques"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Data analysis
library(tidyverse)     

# Modeling
library(tidymodels)
library(caret)

# Visualization
library(naniar)
library(visdat)
library(ComplexHeatmap)


training <- read_csv("train.csv")
testing <- read_csv("test.csv")

```



## Missing data
I start by fixing the obvious missing values. The only problematic variable seems to be year built for garage where the garace is missing. One way would be to replace NA with year built for the house itselfe.   
```{r}
# Missing data according to data_description.txt file

training$PoolQC[is.na(training$PoolQC)] <- "no_pool"
training$MiscFeature[is.na(training$MiscFeature)] <- "none"
training$Alley[is.na(training$Alley)] <- "no_alley"
training$Fence[is.na(training$Fence)] <- "no_fence"
training$FireplaceQu[is.na(training$FireplaceQu)] <- "no_fireplace"

#  Basement
training$BsmtFinType2[is.na(training$BsmtFinType2)] <- "no_basement"
training$BsmtFinType1[is.na(training$BsmtFinType1)] <- "no_basement"
training$BsmtExposure[is.na(training$BsmtExposure)] <- "no_basement"
training$BsmtQual[is.na(training$BsmtQual)] <- "no_basement"
training$BsmtCond[is.na(training$BsmtCond)] <- "no_basement"

# Garage
training$GarageType[is.na(training$GarageType)] <- "no_garage"
training$GarageFinish[is.na(training$GarageFinish)] <- "no_garage"
training$GarageQual[is.na(training$GarageQual)] <- "no_garage"
training$GarageCond[is.na(training$GarageCond)] <- "no_garage"


df_missing <- map_df(training, function(x) mean(is.na(x))) %>%
        gather()

ggplot(filter(df_missing, value > 0),
       aes(x = fct_reorder(key, value),
           y = value)) +
        geom_bar(stat = "identity") +
        coord_flip()

```


### Further investigation into missing values
```{r}
convert_missing <- function(x) ifelse(is.na(x), 0, 1)

house_missing <- apply(training, 2, convert_missing)

Heatmap(house_missing,
        name = "Missing",
        column_title = "Predictors",
        row_title = "Samles",
        col = c("black", "lightgrey"),
        show_heatmap_legend = FALSE,
        row_names_gp = gpar(fontsize = 0)) # text size for row names


gg_miss_upset(training)


```


### LotFrontage
LotFrontage referce to *Linear feet of street connected to property*. When compared to SalePrice there doesn't seems to be that much of a difference in sale price whether or not the LotFrontage is missing. It could be that the actual value is 0 but we dont' know for sure. So we can impute the missing values using KNN or Bagging and compare it to imputing the missing values with zero.  

#### Numeric variables and missingness of LotFrontage
No clear pattern.  
```{r}

df_numeric <- training %>% 
        select_if(is.numeric) %>% 
        mutate(LotFrontage_cat = ifelse(is.na(LotFrontage), "missing", "no_missing")) %>% 
        select(-Id,- LotFrontage) %>% 
        select(LotFrontage_cat, everything()) %>% 
        gather("key", "value", 2:37)

ggplot(df_numeric,
       aes(x = LotFrontage_cat,
           y = value,
           col = LotFrontage_cat)) + 
        geom_violin() +
        facet_wrap(~key, ncol = 6, scales = "free_y")

```

#### Categorical variables and connection to LotFrontage

```{r}
df_categorical <- training %>% 
        select_if(is.character)

df_lot <- training %>% 
        select(Id, LotFrontage)

df_categorical <- bind_cols(df_lot, df_categorical) %>% select(-Id)

df_categorical <- df_categorical %>% 
        gather("key", "value", 2:44) %>% 
        group_by(key, value) %>% 
        summarise(missing = mean(is.na(LotFrontage)),
                  fjoldi = n()) %>% 
        arrange(desc(missing))

head(df_categorical, 10)

# ggplot(df_categorical,
#        aes(x = value,
#            y = missing,
#            fill = value)) +
#         geom_bar(stat = "identity") +
#         facet_wrap(~key) + 
#         theme(legend.position = "none")

```


### GarageYrBlt
Problematic variable. It is numeric but contains NA's for properties with no garage. Could impute with the year the house was built. First I'll check if this variable is important or not using random forest.
```{r}
# I have to change all character variables to factor. Got an error: Error in gower_work....

train_no_garageyrblt <- training[!is.na(training$GarageYrBlt), ] %>% select(-Id) %>% 
        mutate_if(is.character, as.factor) %>% 
        janitor::clean_names()

ames_rf_importance <- recipe(sale_price ~ ., data = train_no_garageyrblt) %>% 
        step_knnimpute(all_predictors(), neighbors = 5) %>% 
        prep(train_no_garageyrblt) %>% 
        bake(train_no_garageyrblt)

n_features <- length(setdiff(names(train_no_garageyrblt), "sale_price"))


rf_model_importance <- ranger::ranger(sale_price ~ .,
                                      data = ames_rf_importance,
                                      mtry = floor(n_features / 3),
                                      respect.unordered.factors = "order",
                                      importance = "impurity")


```

#### Is GarageYrBlt important?
```{r}
df_rf_importance <- as.data.frame(rf_model_importance$variable.importance) %>% 
        rownames_to_column(var = "variable") %>% 
        as_tibble()

colnames(df_rf_importance) <- c("variable", "importance")
df_rf_importance <- df_rf_importance %>% arrange(desc(importance))

ggplot(df_rf_importance,
       aes(x = fct_reorder(variable, -importance),
           y = importance)) +
        geom_bar(stat = "identity") + 
        coord_flip()


rank_garagyrb <- which(df_rf_importance$variable == "garage_yr_blt")
yrblt_garageyrb <- mean(train_no_garageyrblt$year_built == train_no_garageyrblt$garage_yr_blt)

```

The year when the garage is built ranks number `r rank_garagyrb` of `r n_features` variables in the dataset. I could replace the garage_yr_blt with year_built. `r yrblt_garageyrb` of garages were built the same year as the house. So I'm not 100% sure if I should impute garage_yr_built with year_built.

```{r}

ggplot(train_no_garageyrblt,
       aes(x = year_built,
           y = garage_yr_blt)) + 
        geom_point()

```



### Recipe
```{r}
# ames_recipe_knn <- recipe(SalePrice ~ ., data = training) %>% 
#         step_log(all_outcomes()) %>% 
#         step_knnimpute()
```

